Hydra Battles and Cie

### 2021-08-21 doc/part-hydras.tex
conversion to Alectryon (chapter 5, to be continued)

    So, a  natural question is `` Does there exist any strictly decreasing variant mapping
    type \texttt{Hydra} into some interval $[0,\alpha[$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''. The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.

@@

@@

\index{hydras}{Projects}
\begin{project}
Many lemmas presented in this chapter were stated and proved before the introduction of 
the type class \texttt{ON} of ordinal notations, and in particular its  instance \texttt{Epsilon0}.
Thus definitions and lemmas refer to the type \texttt{T1} of possibly not well-formed terms.
This should be fixed in  a future version.
\end{project}

@@

\label{lemma:canonS_LT}
\begin{Coqsrc}
Lemma canonS_LT i alpha :
  nf alpha -> alpha <> zero ->
  canon alpha (S i) t1<  alpha.
\end{Coqsrc}

@@

\vspace{4pt}
\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}

\begin{Coqsrc}
Lemma canonS_limit_strong (lambda : T1) : 
     nf lambda ->
     limitb lambda  ->
     forall beta, beta t1< lambda ->
                  {i:nat | beta t1< canon lambda (S i)}.

Proof.
  transfinite_induction_LT lambda.
  (* ... *)
Defined.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma canonS_limit_lub (lambda : T1) :
  nf lambda -> limitb lambda  ->
  strict_lub (canonS lambda) lambda.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition transition_S i : relation T1 :=
fun alpha beta =>
    alpha <> zero /\ beta = canonS alpha (S i).

Definition transition i : relation T1 :=
  match i with 0 => fun _ _ => False | S j => transition_S j end.

Inductive path_to (beta: T1) : list nat -> T1 -> Prop :=
  path_to_1 : forall (i:nat) alpha , 
    i <> 0 ->
    transition i alpha beta ->
    path_to beta (i::nil) alpha
| path_to_cons : forall i alpha s gamma,
    i <> 0 ->
    transition i alpha gamma ->
    path_to beta  s gamma ->
    path_to beta  (i::s) alpha.
\end{Coqsrc}

\begin{remark}
The definition above is parameterized with the \emph{destination} of the path and indexed by the origin, hence the name \texttt{path\_to}. The rationale behind this choice is a personal preference of the developer  for the kind of eliminators generated by \coq{} in this case. The symmetric option could have been also considered (see also Remark~\vref{remark:transitive-closure}).
\end{remark}

\begin{remark}
In the present version of our library, we use a variant \texttt{path\_toS} of
\texttt{path\_to}, where the proposition
(\texttt{path\_toS $\beta$ $s$ $\alpha$}) is equivalent to
(\texttt{path\_to $\beta$ (shift $s$) $\alpha$}). This variant is scheduled to be deprecated.
\end{remark}

@@

@@

\begin{Coqsrc}
Lemma path_to_LT beta s alpha :
  path_to beta s alpha -> nf alpha -> beta t1< alpha.
\end{Coqsrc}

@@

The Gaia project ~\url{https://github.com/coq-community/gaia} maintains Grimm's  theory of ordinals as part of coq-community on GitHub. Integration
of the present ordinal theory with Gaia, i.e., relating the different notions of ordinals
and transferring relevant results, is an interesting project.
First experiments in that direction are developped in
\href{https://github.com/coq-community/hydra-battles/blob/master/theories/gaia/T1Bridge.v}{theories/gaia/T1Bridge.v}.

>>>>

    So, a  natural question is `` Does there exist any strictly decreasing variant mapping
    type \texttt{Hydra} into some interval $[0,\alpha)$ (where $\alpha <\epsilon_0$) for proving the termination of all hydra battles''. The next chapter is dedicated to a formal proof that there exists no such $\alpha$, even if we consider a restriction to the set of ``standard'' battles.

@@

In the present state of this library, the following specializations of \texttt{canon} are still used in some proofs or lemma statements.

\input{movies/snippets/Canon/CanonS0}

@@

% \index{hydras}{Projects}
% \begin{project}
% Many lemmas presented in this chapter were stated and proved before the introduction of 
% the type class \texttt{ON} of ordinal notations, and in particular its  instance \texttt{Epsilon0}.
% Thus definitions and lemmas refer to the type \texttt{T1} of possibly not well-formed terms.
% This should be fixed in  a future version.
% \end{project}

@@

\label{lemma:canonS_LT}

\input{movies/snippets/Canon/canonSLT}

@@

\emph{From Module~\href{../theories/html/hydras.Epsilon0.Canon.html\#canonS_limit_strong}{Epsilon0.Canon}}

\input{movies/snippets/Canon/canonSLimitStrong}

@@

\input{movies/snippets/Canon/canonSLimitLub}

@@

\input{movies/snippets/Paths/transitionDefs}

\input{movies/snippets/Paths/pathDef}

\begin{remark}
In the present version of our library, we use a variant \texttt{path\_toS} of
\texttt{path\_to}, where the proposition
(\texttt{path\_toS $\beta$ $s$ $\alpha$}) is equivalent to
(\texttt{path\_to $\beta$ (List.map S $s$) $\alpha$}). This variant is scheduled to be deprecated.
\end{remark}

@@

\input{movies/snippets/Paths/LTPathTo}

\input{movies/snippets/Paths/pathToLT}

@@

@@

The Gaia project ~\url{https://github.com/coq-community/gaia} maintains Grimm's  theory of ordinals as part of coq-community on GitHub. Integration
of the present ordinal theory with Gaia, i.e., relating the different notions of ordinals
and transferring relevant results, is an interesting project.
First experiments in that direction are developped in
the \href{https://github.com/coq-community/hydra-battles/blob/master/theories/gaia/}{theories/gaia/} directory.
